---
title: "Processing and visualising the NOAA Dataset of Significant Earthquakes"
author: "maianhdang"
date: "September 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Info

[![Travis-CI Build Status](https://travis-ci.org/maianhdang/noaa.visualise.svg?branch=master)](https://travis-ci.org/maianhdang/noaa.visualise)

This is `noaa.visualise` package working with the NOAA Significant Earthquakes dataset 
to support the process of visualising and gaining information. 

The packages includes the below groups:

Group                 | Functions
----------------------| ----------------------------
Cleaning the raw data | eq_clean_data(), eq_location_clean(), eq_country_filter()
Visualisation tools   | geom_timeline(), geom_timeline_lable(), stat_timeline(), theme_timeline()
Mapping tools         | eq_map(), eq_create_label()

## Included dataset


The [NOAA Significant Earhquake Dataset](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) contains the information of 5,933 earthquakes over the period of 4,000 years.
**Cite:** National Geophysical Data Center / World Data Service (NGDC/WDS): Significant Earthquake Database. National Geophysical Data Center, NOAA. doi:10.7289/V5TD9V7K

## Install the package

```{r visual_1}
devtools::install_github('maianhdang/noaa.visualise', build_vignettes = TRUE)
library(noaa.visualise)
```

## Further information about usage
The vignettes is included in the package. Call the vignettes by:
`vignette('noaa_instruct', package = 'noaa.visualise')`

