---
title: "NOAA Visualization Package"
author: "Anh Dang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is `noaa.visualise` package, working with the NOAA Significant Earthquakes dataset 
to support the process of visualising and gaining information, including functions to: 

- Cleaning and filtering data
- Plot the significant earthquakes in NOAA datasets by timeline
- Plot the significant earthquakes in NOAA datasets by longitude and latitude

## Package Structure

The packages includes the below groups:
  
Group                 | Functions
----------------------| ----------------------------
Cleaning the raw data | eq_clean_data(), eq_location_clean(), eq_country_filter()
Visualisation tools   | geom_timeline(), geom_timeline_lable(), stat_timeline(), theme_timeline()
Mapping tools         | eq_map(), eq_create_label()

## Cleaning the raw data

To make the data appropriate to visualize by the tools provided in the package. You should clean the data:

```{r clean_data}
data("raw_data")
data <- noaa.visualise::eq_clean_data(raw_data)
data <- noaa.visualise::eq_location_clean(data) ## to create the text annotation
input_data <- noaa.visualise::eq_country_filter(data, countries = c("MEXICO", "IRAN"),
                                            xmin = as.Date("1000-01-01"),
                                            xmax = as.Date("2010-01-01"))
```

## geom_timeline and geom_timeline_label

These geom are used altogether with the `gglot2`. You can choose the visualized timezone by `xmin` and `xmax` parameters. Mutiple information could be set by `size` and `color` (similar to other ggplot2 geoms).
`geom_timeline_label` would add the `LOCATION_NAME` to the timeline, the number of label would be assigned to `n_max` largest earthquakes.

```{r visual_1}
data("raw_data")
data <- noaa.visualise::eq_clean_data(raw_data)
data <- noaa.visualise::eq_location_clean(data) ## to create the text annotation
input_data <- noaa.visualise::eq_country_filter(data, countries = c("MEXICO", "IRAN"),
                                            xmin = as.Date("1000-01-01"),
                                            xmax = as.Date("2010-01-01"))

ggplot2::ggplot(input_data, ggplot2::aes (x = date, y = COUNTRY, color = as.numeric(DEATHS), size = as.numeric(EQ_PRIMARY))) +

## geom_timeline  
noaa.visualise::geom_timeline(ggplot2::aes(xmin = as.Date("1900-01-01"), xmax =  as.Date("1925-01-01"))) +
ggplot2::labs(size = "Richter Scale Values", color = "Number of Deaths" )+

## geom_timeline_label  
noaa.visualise::geom_timeline_label(ggplot2::aes(label = LOCATION_NAME, 
                          xmin = as.Date("1900-01-01"), xmax = as.Date("1925-01-01")), 
                          n_max = 5) + 
  
noaa.visualise::theme_timeline()
```

## Visualize earthquakes in the map
The earthquakes could be visualized on the geographical maps, with specific pop-up information, with the `eq_map`. The `eq_create_label` could be used together to create more meaningful popup texts.

```{r map_1}
library(magrittr)
data("raw_data")
data <- noaa.visualise::eq_clean_data(raw_data)
input_data <- dplyr::filter(data, COUNTRY == "MEXICO" &
                                          lubridate::year(date) >= 2000)%>%
  dplyr::mutate(popup_text = noaa.visualise::eq_create_label(.)) ## create popup_text

noaa.visualise::eq_map(input_data, annot_col = "popup_text") ## annot_col set as popup_text
          
```
